Deterministic Scheduling of Periodic Messages in a Network

### 2017-07-19 infocom/infocom.tex
11 pages

      A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a sequence  ${\cal M}=(m_0, \ldots ,m_{n-1})$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $0$ and $P-1$. Moreover, \emph{no pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.

      As an example of a $(2,1)$-periodic assignment, let us consider a routed network 
      where all pairs of routes intersect at a different arc. It is easy to design such a network and an example is given in Fig.~\ref{fig:example}. We set $\tau = 1$ and the weights are chosen so that if $r_{i}$ and $r_{j}$ have $v$ as first common vertex then we have $\lambda(v,r_{i}) - \lambda(v,r_{j})=1$. There is a $(2,1)$-periodic assignment by setting all $m_{i}$ to $0$.

      \begin{figure}[ht]
      \begin{center}
          \scalebox{0.5}{

\begin{tikzpicture}

\tikzset{
  LabelStyle/.style = { rectangle, rounded corners, draw,
                       font = \bfseries },
  EdgeStyle/.append style = {->} }
  \SetGraphUnit{5}
  \node[draw,circle] (s3) at (4, 2) {$s_2$}; 
  \node[draw,circle] (s2) at (0, 4) {$s_1$}; 
  \node[draw,circle] (s1) at (0, 6) {$s_0$};

  \node[draw,circle] (t3) at (14, 7) {$t_2$}; 
  \node[draw,circle] (t2) at (14, 4) {$t_1$}; 
  \node[draw,circle] (t1) at (10, 2) {$t_0$};

  \SetVertexNoLabel
  \Vertex[x=2,y=5]{A}
  \Vertex[x=4,y=5]{B}
  \Vertex[x=10,y=5]{C}
  \Vertex[x=12,y=5]{D}
  \Vertex[x=6,y=3]{E}
  \Vertex[x=8,y=3]{F}
  \tikzset{
  EdgeStyle/.append style = {green} }
  \Edge[label = 2](s2)(A)
  \Edge[label = 1](A)(B)
  \Edge[label = 2](B)(C)
  \Edge[label = 1](C)(D)
  \Edge[label = 1](D)(t2)

   \tikzset{
  EdgeStyle/.append style = {red} }
  \Edge[label = 2](s3)(E)
  \Edge[label = 1](E)(F)
  \Edge[label = 1](F)(C)
  \Edge[label = 1](C)(D)
  \Edge[label = 1](D)(t3) 
     \tikzset{
  EdgeStyle/.append style = {blue} }
  \Edge[label = 1](s1)(A)
  \Edge[label = 1](A)(B)
  \Edge[label = 1](B)(E)
  \Edge[label = 1](E)(F)
  \Edge[label = 1](F)(t1)

\end{tikzpicture}

}
     \end{center}
       \caption{A routed network with $(0,0,0)$ as a $(2,1)$-periodic assignment}
       \label{fig:example}
      \end{figure}

      \subsection{Periodic route assignment}\label{nonmonotone}

    We want to find an assignment which allows to send periodic messages from sources to targets
    without collisions. We introduce the following associated decision problem, useful for hardness proofs.

      \noindent {\bf  Periodic Routes Assignment (PRA)}

      \noindent {\bf Input:} a routed network $(G,\cal R)$, an integer $\tau$ and an integer $P$.

      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic assignment of $(G,\cal R)$ ?

      We will prove in Sec.~\ref{sec:complexity} that the problem PRA is $\NP$-complete, even in restricted settings.
      In fact, approximating the smallest value of $P$ for which there is a $(P,\tau)$-periodic assignment is already hard.

      An unusual property of assignment is that given a routed network, we may have a $(P,\tau)$-periodic assignment but no
      $(P',\tau)$-periodic assignment with $P' > P$: the existence of an assignment is not monotone with regard to $P$.

	\begin{lemma} \label{lemma:monotonic}
	 For any odd $P$, there is a routed network such that there is $(2,1)$-periodic assignment but no $(P,1)$-periodic assignment.
	\end{lemma}
\begin{proof}

      Consider the routed network $(G,{\cal R})$ given in the previous subsection. 
      We change the weights so that for $v$, the first vertex which belongs to $r_i$ and $r_j$,
      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $n$, the number of routes in ${\cal R_{\cal C}}$. In such a graph, there is no $(P,\tau)$-periodic assignment, since the problem reduces to finding a $P$-coloring in a complete graph with $n > P$ vertices, the colors being the offsets of the routes.\\
      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) \mod 2 = 1$ . Therefore $(0,\dots,0)$ is a $(2,1)$-periodic assignment of ${\cal R}$.      
\end{proof}

         We now give a new interpretation of a $(P,\tau)$-periodic assignment of a $(G,{\cal R})$ symmetric routed network, so that it represents the sending of a message and of its answer.
	This assignment represents the following process: first a message is sent at $u$, through the route $r \in F$, at time $m_r$.

      As a consequence of the $\NP$-hardness of PRA, we show in the next subsection that this problem is $\NP$-hard. 
      In Sec.~\ref{sec:PALL} we will study heuristics which try to solve the search version of PALL (computing an assignment), also denoted by PALL for simplicity.

\section{Hardness of PRA}
  \label{sec:complexity}

 In this section we always assume that the size of a message $\tau$ is equal to one. 
 We will prove the hardness of PRA and PALL for $\tau =1$ which implies the hardness of the general problems. 
Consider an instance of the problem PRA, i.e., a routed network $(G,\cal{R})$ and a period $P$.
The {\bf conflict depth} of a route is the number of arcs of the route which also belong to other routes.
The conflict depth of a routed network $(G,\cal{R})$ is the maximum of the conflict depth of its routes.
The {\bf load} of a routed network is the maximal number of routes sharing the same arc.
Remark that a $(P,1)$-periodic assignment must satisfy that $P$ is larger or equal to the load.

We give two alternate proofs that PRA is $\NP$-complete.
The first proof works already for conflict depth two. Remark that for conflict depth one,
the graph can be seen as a set of disjoint pair of routes, on which PRA and PALL can be solved in linear time. 
 The second proof reduces the problem to graph coloring and implies inapproximability when one tries to find the smallest possible $P$. \\
 Finally, it is easy to see that PRA is easy on trees and it may be interesting to study its complexity on 
 bounded treewidth networks, since it is a common property of real networks~\cite{de2011treewidth}.

 \begin{theorem}
Problem PRA is $\NP$-complete on the class of routed networks with conflict depth two.
\end{theorem}
 \begin{proof}
 Problem $PRA$ is in $\NP$ since given an offset for each route in an assignment, it is easy to check in linear time in the number of arcs whether there are collisions.

  Let $H=(V,E)$ be an undirected graph and let $d$ be its maximal degree. We consider the problem to determine whether $H$ is arc-colorable
  with $d$ or $d+1$ colors. The arc coloring problem is $\NP$-hard~\cite{holyer1981np} and we reduce it to PRA to prove its $\NP$-hardness. To do that, we define from $H$ a routed network $(G,{\cal R})$ as follows.
  The vertices of $G$ are $v_1, v_2$ for each $v$ in $V$ and $s_{u,v}, t_{u,v}$ for each $(u,v) \in E$.
  For each arc $(u,v) \in E$, there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$.

   To define ${\cal R}$ an arbitrary orientation of each arc $(u,v)$ is chosen. 
   Then for each arc $(u,v)$ there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$.
  All these arcs are of weight $0$. The set of arcs of G is the union between all the arcs of the previously defined routes.

  Observe that the existence of a $d$-coloring of $H$ is equivalent to the existence of a $(d,1)$-periodic assignment
  of $(G,{\cal R})$. Indeed, a $d$-coloring of $H$ can be seen as a labeling of its arcs by the integers
  in $\{0,\dots,d-1\}$ and we have a bijection between $d$-colorings of $H$ and offsets of the routes of ${\cal R}$.
  By construction, the constraint of having no collision between the routes is equivalent to the fact that no two adjacent arcs have the same color. Therefore we have reduced arc coloring to PRA by a polynomial time transformation which concludes the proof. 
 \end{proof}

 Remark that we have used zero weights in the proof. If we ask the weights to be strictly positive, which makes sense in our model since
they represent the delay of physical links, it is easy to adapt the proof. We just have to set them so that in any route the weight at $u_1$ is equal to $d$ and thus equal to $0$ modulo $d$. We now lift this hardness result to the problem PALL.

\begin{corollary}
Problem PALL is $\NP$-complete on the class of routed networks of conflict depth two.
\end{corollary}
\begin{proof}
 We consider $((G,{\cal R}),P,\tau)$ an instance of $PRA$. We assume that no vertex is the first of some route and the last of another one. Remark that this condition is satisfied in the previous proof, which makes the problem $PRA$ restricted to this kind of instance $\NP$-complete. 
 Let us define $T_{max} = 2 \times \max_{r \in {\cal R}} \lambda(r) + P$. We define $(G',{\cal R}')$ a symmetric routed network from $(G,{\cal R})$ where for every route we add a symmetric route with new arcs of opposite orientation and the same weights.
 We prove that the instance $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL if and only if $((G,{\cal R}),P,\tau)$ is in $PRA$.
 If $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL, then a $(P,\tau)$-assignment of $(G',{\cal R'})$ restricted to $\R$ is a $(P,\tau)$-assignment of $((G,{\cal R})$ since they cannot be any collision between routes of ${\cal R}$.

 Assume now that $((G,{\cal R}),P,\tau)$ is in $PRA$. First remark that the waiting time of each route is by definition less than $P$ and thus the maximal process time is always less than $T_{max}$, therefore the constraint on $T_{max}$ is satisfied. Moreover a $(P,\tau)$-assignment of $(G,{\cal R})$ can be extended into a $(P,\tau)$-assignment of $(G',{\cal R'})$ in the following way. For each route $r \in \cal{R}$ of last vertex $v$, the time at which the message arrives is $t(v,r)$, then we choose as offset for $\rho(r) \in \cal{R}'$ $-t(v,r) \mod P$. The symmetry ensures that each new route $\rho(r)$ in ${\cal R'}$ uses exactly the same times slot as $r$ one each of its node and thus avoid collisions.
\end{proof}

Let MIN-PRA be the problem, given a routed network and an assignment, to find the minimal period $P$ such that there is a $P$-periodic assignment.

\begin{theorem}\label{th:inapprox}
If $\P \neq \NP$, problem MIN-PRA on the classes of routed networks of load two cannot be approximated in polynomial time within a factor $n^{1-o(1)}$ where $n$ is the number of routes.
\end{theorem}

\begin{proof}
 We reduce graph coloring to PRA. Let $H$ be a graph instance of the $k$-coloring problem. 
 We define ${\cal R}$ in the following way: for each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$.
 Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. Note that it is easy to build a graph with such routes as in Fig.~\ref{fig:reduction}.

 Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
 by converting an offset of a route into a color of a vertex and reciprocally. Therefore if we can approximate the minimum value of $P$ within some factor, we could approximate the minimal number of colors needed to color a graph within the same factor. The proof follows from the hardness of approximability of finding a minimal coloring~\cite{zuckerman2006linear}.
\end{proof}

In particular, this reduction shows that even with small maximal load, the minimal period can be large.

    \begin{figure}[ht]

    \hspace{0.5cm}\scalebox{0.35}{
    \begin{tikzpicture}
    \tikzset{
      LabelStyle/.style = { rectangle, rounded corners, draw,
			  font = \bfseries },
      EdgeStyle/.append style = {->} }
      \SetGraphUnit{5}

      \node[draw,circle] (s3) at (4, 2) {$s_2$}; 
      \node[draw,circle] (s2) at (0, 4) {$s_1$}; 
      \node[draw,circle] (s1) at (0, 6) {$s_0$};

      \node[draw,circle] (t3) at (12, 3) {$t_2$}; 
      \node[draw,circle] (t2) at (14, 4) {$t_1$}; 
      \node[draw,circle] (t1) at (10, 2) {$t_0$};

      \tikzstyle{VertexStyle}=[shape = circle, draw, minimum size = 20pt]
	\tikzset{
      VertexStyle/.append style = {blue} }
	\Vertex[x=-8,y=3]{1}
	      \tikzset{
      VertexStyle/.append style = {green} }
	  \Vertex[x=-7,y=5]{2}

	    \tikzset{
      VertexStyle/.append style = {red} }
	  \Vertex[x=-6,y=4]{3}
		\tikzset{
      VertexStyle/.append style = {black} }

      \SetVertexNoLabel
      \Vertex[x=2,y=5]{A}
      \Vertex[x=4,y=5]{B}
      \Vertex[x=10,y=5]{C}
      \Vertex[x=12,y=5]{D}
      \Vertex[x=6,y=3]{E}
      \Vertex[x=8,y=3]{F}
      \tikzset{
      EdgeStyle/.append style = {green} }
      \Edge(s2)(A)
      \Edge(A)(B)
      \Edge(B)(C)
      \Edge(C)(D)
      \Edge(D)(t2)

      \tikzset{
      EdgeStyle/.append style = {red} }
      \Edge(s3)(E)
      \Edge(E)(F)
      \Edge(F)(t3) 
	\tikzset{
      EdgeStyle/.append style = {blue} }
      \Edge(s1)(A)
      \Edge(A)(B)
      \Edge(B)(E)
      \Edge(E)(F)
      \Edge(F)(t1)

	\tikzset{
      EdgeStyle/.append style = {black,-} }

      \Edge(1)(2)
      \Edge(1)(3)
    \node (1) at (-3,4){\Huge $\rightarrow$};
%     
%     \node (2) at (-7,0){\Huge H};
%     \node (3) at (10,0){\Huge G};
    \end{tikzpicture}
    }
    \caption{Reduction from k-coloring to MIN-PRA}
    \label{fig:reduction}
    \end{figure}

@@

@@

  In this subsection, we deal with a simpler version of the problem PALL.
  We ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$} and we call this restricted problem {\bf Periodic Assignment for Zero Latency} or PAZL. We consider PAZL since it is simpler to get theoretical results and good algorithms 
  than for PALL. Moreover, as we show in our experimentations of Sec.\ref{sec:exp_PAZL}, this problem can often be solved positively (albeit less often than the general problem). Finally, a solution to PAZL is simpler to implement in real telecommunication networks, since we do not need to implement any buffering at all.

  When the waiting times are zero, $MT({\cal M})$ is equal to twice the length of the longest route, thus $T_{max}$ is not relevant anymore. For a route $r$, choosing the offset $m_r$ also sets the offset of the route $\rho(r)$ to $m_{\rho(r)} = m_{r} + \lambda(r) \mod P$.
  Consider an assignment $(m_0,\dots,m_{n-1})$ of a star routed network and let $m_i'= m_{i} - t(c_s,r_i) \mod P$.
  Then $\{m_0',\dots,m_{n-1}'\}$ is an assignment of the same star routed network where the weights of the arcs from $s_i$ to $c_s$ are $0$ for all $i$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

\begin{figure}[H] 
    \begin{minipage}[c]{0.5\textwidth}
  \centering
      \includegraphics[width=0.5\textwidth]{departs_gp_21000.pdf}

      \caption{Success rate of different orders, $95\%$ load.}
      \end{minipage} \hfill
        \begin{minipage}[c]{0.5\textwidth}
          \includegraphics[width=0.5\textwidth]{departs_gp_25000.pdf}
      \caption{Success rate of different orders, $80\%$ load.}

            \end{minipage} \hfill
                  \label{fig:success30000}
     \end{figure}

@@

    \begin{figure} [H] 
       \begin{center}
      \includegraphics[scale=0.45]{retour_21000.pdf}
      \end{center}
      \caption{Success rate of GD, MLS and PMLS for $P = 21000$}
     \label{fig:success21000}
     \end{figure}

@@

         \begin{figure}[H] 
       \begin{center}
   \begin{tabular}{|c|c|c|c|c|c|c|}
    \hline
    Number of random orders & $1$ & $10$ & $100$& $1 000$& $10 000$&$100 000$\\
    \hline
    Percentage of success of GD & $0.6$ &$7.1$&$33.4$&$76.4$&$90.0$&$90.0$\\
    \hline
    Percentage of success of PMLS & $41,1$ &$82,5$&$94,5$&$96,2$&$97,2$&$97,2$\\
    \hline
      \end{tabular}
      \end{center}
   \caption{Impact of the number of random sending orders}
     \end{figure}

@@

       \begin{center}
      \includegraphics[scale=0.4]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \caption{Needed margin with respect to the period for a stochastic policy}
      \label{fig:sto}
     \end{figure}

>>>>

@@

%      A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a sequence  ${\cal M}=(m_0, \ldots ,m_{n-1})$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ represents the time at which a message is emitted at the first vertex of the route $r_{i}$. We will always consider that the $m_{i}$ are between $0$ and $P-1$. Moreover, \emph{no pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.
%      As an example of a $(2,1)$-periodic assignment, let us consider a routed network 
%      where all pairs of routes intersect at a different arc. It is easy to design such a network and an example is given in Fig.~\ref{fig:example}. We set $\tau = 1$ and the weights are chosen so that if $r_{i}$ and $r_{j}$ have $v$ as first common vertex then we have $\lambda(v,r_{i}) - \lambda(v,r_{j})=1$. There is a $(2,1)$-periodic assignment by setting all $m_{i}$ to $0$.
%      \begin{figure}[ht]
%      \begin{center}
%          \scalebox{0.5}{
%\begin{tikzpicture}
%\tikzset{
%  LabelStyle/.style = { rectangle, rounded corners, draw,
%                       font = \bfseries },
%  EdgeStyle/.append style = {->} }
%  \SetGraphUnit{5}
%  \node[draw,circle] (s3) at (4, 2) {$s_2$}; 
%  \node[draw,circle] (s2) at (0, 4) {$s_1$}; 
%  \node[draw,circle] (s1) at (0, 6) {$s_0$}; 
%  \node[draw,circle] (t3) at (14, 7) {$t_2$}; 
%  \node[draw,circle] (t2) at (14, 4) {$t_1$}; 
%  \node[draw,circle] (t1) at (10, 2) {$t_0$}; 
%  \SetVertexNoLabel
%  \Vertex[x=2,y=5]{A}
%  \Vertex[x=4,y=5]{B}
%  \Vertex[x=10,y=5]{C}
%  \Vertex[x=12,y=5]{D}
%  \Vertex[x=6,y=3]{E}
%  \Vertex[x=8,y=3]{F}
%  \tikzset{
%  EdgeStyle/.append style = {green} }
%  \Edge[label = 2](s2)(A)
%  \Edge[label = 1](A)(B)
%  \Edge[label = 2](B)(C)
%  \Edge[label = 1](C)(D)
%  \Edge[label = 1](D)(t2)
%   \tikzset{
%  EdgeStyle/.append style = {red} }
%  \Edge[label = 2](s3)(E)
%  \Edge[label = 1](E)(F)
%  \Edge[label = 1](F)(C)
%  \Edge[label = 1](C)(D)
%  \Edge[label = 1](D)(t3) 
%     \tikzset{
%  EdgeStyle/.append style = {blue} }
%  \Edge[label = 1](s1)(A)
%  \Edge[label = 1](A)(B)
%  \Edge[label = 1](B)(E)
%  \Edge[label = 1](E)(F)
%  \Edge[label = 1](F)(t1)
%\end{tikzpicture}
%}
%     \end{center}
%       \caption{A routed network with $(0,0,0)$ as a $(2,1)$-periodic assignment}
%       \label{fig:example}
%      \end{figure}
%      \subsection{Periodic route assignment}\label{nonmonotone}
%    We want to find an assignment which allows to send periodic messages from sources to targets
%    without collisions. We introduce the following associated decision problem, useful for hardness proofs.
%      \noindent {\bf  Periodic Routes Assignment (PRA)} 
%      \noindent {\bf Input:} a routed network $(G,\cal R)$, an integer $\tau$ and an integer $P$.
%      \noindent {\bf Question:} does there exist a $(P,\tau)$-periodic assignment of $(G,\cal R)$ ?
%      We will prove in Sec.~\ref{sec:complexity} that the problem PRA is $\NP$-complete, even in restricted settings.
%      In fact, approximating the smallest value of $P$ for which there is a $(P,\tau)$-periodic assignment is already hard.
%   An unusual property of assignment is that given a routed network, we may have a $(P,\tau)$-periodic assignment but no
%	\begin{lemma} \label{lemma:monotonic}
%	 For any odd $P$, there is a routed network such that there is $(2,1)$-periodic assignment but no $(P,1)$-periodic assignment.
%	\end{lemma}
%\begin{proof}
%      Consider the routed network $(G,{\cal R})$ given in the previous subsection. 
%      We change the weights so that for $v$, the first vertex which belongs to $r_i$ and $r_j$,
%      we have $\lambda(v,r_i) - \lambda(v,r_j)= P$, where $P$ is an odd number smaller than $n$, the number of routes in ${\cal R_{\cal C}}$. In such a graph, there is no $(P,\tau)$-periodic assignment, since the problem reduces to finding a $P$-coloring in a complete graph with $n > P$ vertices, the colors being the offsets of the routes.\\
%      If we consider a period of $2$, for all $i \neq j$, $\lambda(v,r_i) - \lambda(v,r_j) \mod 2 = 1$ . Therefore $(0,\dots,0)$ is a $(2,1)$-periodic assignment of ${\cal R}$.      
%\end{proof}

         A {\bf $(P,\tau)$-periodic assignment} of a routed network $(G,\cal R)$ is a set of two sequence  ${\cal M}=\{(m_0, \ldots ,m_{n-1}),(m_{\rho_0}, \ldots ,m_{\rho_{n-1}})\}$ of $n$ integers that we call {\bf offsets}, with $n$ the number of routes in $\cal R$. The number $m_{i}$ (respectively $m_{\rho_i}$) represents the time at which a message is emitted at the first vertex of the route $r_{i}$ (respectively $r_{\rho_i}$).A $(P,\tau)$-periodic assignment of a $(G,{\cal R})$ symmetric routed network represents
 the sending of a message and of its answer. \emph{No pair of routes must have a collision} in a $(P,\tau)$-periodic assignment.
   This assignment represents the following process: first a message is sent at $u$, through the route $r \in F$, at time $m_r$.

      \todo{mettre le lemme sur la monotonie sans preuve?}

     We show in the long version of this article that this problem is $\NP$-hard. 
      In Sec.~\ref{sec:PALL} we will study heuristics which try to solve the search version of PALL (computing an assignment), also denoted by PALL for simplicity.

One can consider a simpler version of PALL.
       We ask for a $(P,\tau)$-periodic assignment {\bf with all waiting times equal to $0$} and we call this restricted problem {\bf Periodic Assignment for Zero Latency} or PAZL. We consider PAZL since it is simpler to get theoretical results and good algorithms 
  than for PALL. Moreover, as we show in our experimentations of Sec.\ref{sec:exp_PAZL}, this problem can often be solved positively (albeit less often than the general problem). Finally, a solution to PAZL is simpler to implement in real telecommunication networks, since we do not need to implement any buffering at all.
%\section{Hardness of PRA}
%  \label{sec:complexity}
% In this section we always assume that the size of a message $\tau$ is equal to one. 
% We will prove the hardness of PRA and PALL for $\tau =1$ which implies the hardness of the general problems. 
%Consider an instance of the problem PRA, i.e., a routed network $(G,\cal{R})$ and a period $P$.
%The {\bf conflict depth} of a route is the number of arcs of the route which also belong to other routes.
%The conflict depth of a routed network $(G,\cal{R})$ is the maximum of the conflict depth of its routes.
%The {\bf load} of a routed network is the maximal number of routes sharing the same arc.
%Remark that a $(P,1)$-periodic assignment must satisfy that $P$ is larger or equal to the load.
%We give two alternate proofs that PRA is $\NP$-complete.
%The first proof works already for conflict depth two. Remark that for conflict depth one,
%the graph can be seen as a set of disjoint pair of routes, on which PRA and PALL can be solved in linear time. 
% The second proof reduces the problem to graph coloring and implies inapproximability when one tries to find the smallest possible $P$. \\
% Finally, it is easy to see that PRA is easy on trees and it may be interesting to study its complexity on 
% bounded treewidth networks, since it is a common property of real networks~\cite{de2011treewidth}.
% \begin{theorem}
%Problem PRA is $\NP$-complete on the class of routed networks with conflict depth two.
%\end{theorem}
% \begin{proof}
% Problem $PRA$ is in $\NP$ since given an offset for each route in an assignment, it is easy to check in linear time in the number of arcs whether there are collisions.
%  Let $H=(V,E)$ be an undirected graph and let $d$ be its maximal degree. We consider the problem to determine whether $H$ is arc-colorable
%  with $d$ or $d+1$ colors. The arc coloring problem is $\NP$-hard~\cite{holyer1981np} and we reduce it to PRA to prove its $\NP$-hardness. To do that, we define from $H$ a routed network $(G,{\cal R})$ as follows.
%  The vertices of $G$ are $v_1, v_2$ for each $v$ in $V$ and $s_{u,v}, t_{u,v}$ for each $(u,v) \in E$.
%  For each arc $(u,v) \in E$, there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$. 
%   To define ${\cal R}$ an arbitrary orientation of each arc $(u,v)$ is chosen. 
%   Then for each arc $(u,v)$ there is a route $s_{u,v},u_1,u_2,v_1,v_2,t_{u,v}$ in ${\cal R}$.
%  All these arcs are of weight $0$. The set of arcs of G is the union between all the arcs of the previously defined routes.
%  Observe that the existence of a $d$-coloring of $H$ is equivalent to the existence of a $(d,1)$-periodic assignment
%  of $(G,{\cal R})$. Indeed, a $d$-coloring of $H$ can be seen as a labeling of its arcs by the integers
%  in $\{0,\dots,d-1\}$ and we have a bijection between $d$-colorings of $H$ and offsets of the routes of ${\cal R}$.
%  By construction, the constraint of having no collision between the routes is equivalent to the fact that no two adjacent arcs have the same color. Therefore we have reduced arc coloring to PRA by a polynomial time transformation which concludes the proof. 
% \end{proof}
% Remark that we have used zero weights in the proof. If we ask the weights to be strictly positive, which makes sense in our model since
%they represent the delay of physical links, it is easy to adapt the proof. We just have to set them so that in any route the weight at $u_1$ is equal to $d$ and thus equal to $0$ modulo $d$. We now lift this hardness result to the problem PALL.
%\begin{corollary}
%Problem PALL is $\NP$-complete on the class of routed networks of conflict depth two.
%\end{corollary}
%\begin{proof}
% We consider $((G,{\cal R}),P,\tau)$ an instance of $PRA$. We assume that no vertex is the first of some route and the last of another one. Remark that this condition is satisfied in the previous proof, which makes the problem $PRA$ restricted to this kind of instance $\NP$-complete. 
% Let us define $T_{max} = 2 \times \max_{r \in {\cal R}} \lambda(r) + P$. We define $(G',{\cal R}')$ a symmetric routed network from $(G,{\cal R})$ where for every route we add a symmetric route with new arcs of opposite orientation and the same weights.
% We prove that the instance $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL if and only if $((G,{\cal R}),P,\tau)$ is in $PRA$.
% If $((G',{\cal R'}),P,\tau,T_{max})$ is in PALL, then a $(P,\tau)$-assignment of $(G',{\cal R'})$ restricted to $\R$ is a $(P,\tau)$-assignment of $((G,{\cal R})$ since they cannot be any collision between routes of ${\cal R}$.
% Assume now that $((G,{\cal R}),P,\tau)$ is in $PRA$. First remark that the waiting time of each route is by definition less than $P$ and thus the maximal process time is always less than $T_{max}$, therefore the constraint on $T_{max}$ is satisfied. Moreover a $(P,\tau)$-assignment of $(G,{\cal R})$ can be extended into a $(P,\tau)$-assignment of $(G',{\cal R'})$ in the following way. For each route $r \in \cal{R}$ of last vertex $v$, the time at which the message arrives is $t(v,r)$, then we choose as offset for $\rho(r) \in \cal{R}'$ $-t(v,r) \mod P$. The symmetry ensures that each new route $\rho(r)$ in ${\cal R'}$ uses exactly the same times slot as $r$ one each of its node and thus avoid collisions.
%\end{proof}
%Let MIN-PRA be the problem, given a routed network and an assignment, to find the minimal period $P$ such that there is a $P$-periodic assignment. 
%\begin{theorem}\label{th:inapprox}
%If $\P \neq \NP$, problem MIN-PRA on the classes of routed networks of load two cannot be approximated in polynomial time within a factor $n^{1-o(1)}$ where $n$ is the number of routes.
%\end{theorem}
%\begin{proof}
% We reduce graph coloring to PRA. Let $H$ be a graph instance of the $k$-coloring problem. 
% We define ${\cal R}$ in the following way: for each vertex $v$ in $H$, there is a route $r_v$ in ${\cal R}$.
% Two routes $r_v$ and $r_u$ share an arc if and only if $(u,v)$ is an arc in $H$; this arc is the only one shared by these two routes. All arcs are of weight $0$. Note that it is easy to build a graph with such routes as in Fig.~\ref{fig:reduction}.
% Observe that the existence of a $k$-coloring of $H$ is equivalent to the existence of a $(k,1)$-periodic assignment in $G$, 
% by converting an offset of a route into a color of a vertex and reciprocally. Therefore if we can approximate the minimum value of $P$ within some factor, we could approximate the minimal number of colors needed to color a graph within the same factor. The proof follows from the hardness of approximability of finding a minimal coloring~\cite{zuckerman2006linear}.
%\end{proof}
%In particular, this reduction shows that even with small maximal load, the minimal period can be large.
%    \begin{figure}[ht]
%    \hspace{0.5cm}\scalebox{0.35}{
%    \begin{tikzpicture}
%    \tikzset{
%      LabelStyle/.style = { rectangle, rounded corners, draw,
%			  font = \bfseries },
%      EdgeStyle/.append style = {->} }
%      \SetGraphUnit{5}
%      \node[draw,circle] (s3) at (4, 2) {$s_2$}; 
%      \node[draw,circle] (s2) at (0, 4) {$s_1$}; 
%      \node[draw,circle] (s1) at (0, 6) {$s_0$}; 
%      \node[draw,circle] (t3) at (12, 3) {$t_2$}; 
%      \node[draw,circle] (t2) at (14, 4) {$t_1$}; 
%      \node[draw,circle] (t1) at (10, 2) {$t_0$}; 
%      \tikzstyle{VertexStyle}=[shape = circle, draw, minimum size = 20pt]
%	\tikzset{
%      VertexStyle/.append style = {blue} }
%	\Vertex[x=-8,y=3]{1}
%	      \tikzset{
%      VertexStyle/.append style = {green} }
%	  \Vertex[x=-7,y=5]{2}
%	    \tikzset{
%      VertexStyle/.append style = {red} }
%	  \Vertex[x=-6,y=4]{3}
%		\tikzset{
%      VertexStyle/.append style = {black} }
%      \SetVertexNoLabel
%      \Vertex[x=2,y=5]{A}
%      \Vertex[x=4,y=5]{B}
%      \Vertex[x=10,y=5]{C}
%      \Vertex[x=12,y=5]{D}
%      \Vertex[x=6,y=3]{E}
%      \Vertex[x=8,y=3]{F}
%      \tikzset{
%      EdgeStyle/.append style = {green} }
%      \Edge(s2)(A)
%      \Edge(A)(B)
%      \Edge(B)(C)
%      \Edge(C)(D)
%      \Edge(D)(t2)
%      \tikzset{
%      EdgeStyle/.append style = {red} }
%      \Edge(s3)(E)
%      \Edge(E)(F)
%      \Edge(F)(t3) 
%	\tikzset{
%      EdgeStyle/.append style = {blue} }
%      \Edge(s1)(A)
%      \Edge(A)(B)
%      \Edge(B)(E)
%      \Edge(E)(F)
%      \Edge(F)(t1)
%	\tikzset{
%      EdgeStyle/.append style = {black,-} }
%      \Edge(1)(2)
%      \Edge(1)(3)
%    \node (1) at (-3,4){\Huge $\rightarrow$};
%%     
%%     \node (2) at (-7,0){\Huge H};
%%     \node (3) at (10,0){\Huge G};
%    \end{tikzpicture}
%    }
%    \caption{Reduction from k-coloring to MIN-PRA}
%    \label{fig:reduction}
%    \end{figure}

@@

  In this subsection, we deal with the problem PAZL.  
  When the waiting times are zero, $MT({\cal M})$ is equal to twice the length of the longest route, thus $T_{max}$ is not relevant anymore. For a route $r$, choosing the offset $m_r$ also sets the offset of the route $\rho(r)$ to $m_{\rho(r)} = m_{r} + \lambda(r) \mod P$.
  Consider an assignment $(m_0,\dots,m_{n-1})$ of a star routed network and let $m_i'= m_{i} - t(c_s,r_i) \mod P$.
  Then $\{m_0',\dots,m_{n-1}'\}$ is an assignment of the same star routed network where the weights of the arcs from $s_i$ to $c_s$ are $0$ for all $i$. Therefore we assume from now on that the \emph{weight of the arcs from $s_i$ to $c_s$ are all equal to $0$}.

@@

  \centering
     \hspace{-1.5cm} \includegraphics[width=0.4\textwidth]{departs_gp_21000.pdf}

      \caption{Success rate of different orders, $95\%$ load.}
          \end{figure}
\begin{figure}[H] 
          \includegraphics[width=0.5\textwidth]{departs_gp_25000.pdf}
      \caption{Success rate of different orders, $80\%$ load.}
           \label{fig:success30000}
     \end{figure}

@@

    \begin{figure} [H] 
       \begin{center}
      \includegraphics[width=0.5\textwidth]{retour_21000.pdf}
      \end{center}
      \caption{Success rate of GD, MLS and PMLS for $P = 21000$}
     \label{fig:success21000}
     \end{figure}

@@

         \begin{figure}[H] 
       \begin{center}
   \begin{tabularx}{0.5\textwidth}{|c|X|X|X|X|X|X|}
    \hline
    \# orders& $1$ & $10$ & $100$& $1 000$& $10 000$&$100 000$\\
    \hline
    GD & $0.6$ &$7.1$&$33.4$&$76.4$&$90.0$&$90.0$\\
    \hline
  PMLS & $41,1$ &$82,5$&$94,5$&$96,2$&$97,2$&$97,2$\\
    \hline
      \end{tabularx}
      \end{center}
   \caption{Impact of the number of random sending orders}
     \end{figure}

@@

       \begin{center}
      \includegraphics[width = 0.5\textwidth]{stochastic.pdf}
       %\input{figures/stochastic}
      \end{center}
      \caption{Needed margin with respect to the period for a stochastic policy}
      \label{fig:sto}
     \end{figure}
